\documentclass{ximera}[font=12pt]


\title{Prologue}   % Use xmlatex all -s to compile when SERVE refuses to work.
\author{Kelly Stady}
\license{CC: 0}         % replace with an appropriate license, or set it in xmPreamble

\begin{document}

\begin{abstract}

\end{abstract}

\maketitle
\label{xim:prologue}

\section*{Introduction}

This is a collection of prealgebra lessons for college students.  I will continue addings lessons until the course
is complete.    


\section*{Web Browsers}

Ximera works best with \textbf{Chrome}, but \textbf{Firefox} and \textbf{Safari} should work as well. 


\section*{Entering and Checking Your Answers}

How you enter and check your answers will vary depending on the question type.  

\begin{enumerate}
\item \textbf{Multiple Choice}: Click on a multiple choice option, then click on the Check work box or press Enter on your keyboard. 

\item \textbf{Answer Box}:  Enter your answer, then click the question mark or press Enter on your keyboard. 

\item \textbf{Dropdown Menu}:  These have a dash followed by a question mark inside a blue box.  Click on the blue box and select
the correct answer. 

\item \textbf{Select All}: Click on each option that you think is correct, then click on the Check work box or press 
Enter on your Keyboard.
\end{enumerate}

There is a \textbf{Math Editor} that appears at the top of the webpage when you click inside an answer box.  You can type your
answer in the editor and click Ok and it will enter what you typed in the answer box. It's probably not needed for this class.  
  
If you want to delete your work on a page, go to Settings at the top of the page and select Remove your answers.  


\section*{Hints}



\section*{Suggested Web Page Settings}

\begin{enumerate}
\item To move the table of contents out of the way, place the cursor in the lower right corner and dragg it to the left.  That will make 
content a bit easier to read. 

\item Set your browsers zoom to 125\% for easy reading.  
\end{enumerate}

\section*{Troubleshooting}

If any web page acts up, you can press CTRL F5 on your keyboard to reset the page.  You will lose your work, on that page, 
but no harm in getting more practice. 


\section*{Accessibility}

Use the screen reader NVDA or Jaws along with Math Cat to read the content.

\section*{Contact}

If you find any errors or have any accessibilty issues with this content, please email me at Kelly.Stady@tri-c.edu.


\begin{center}
\begin{tikzpicture}[scale=1.5]
  % Define color palette using the requested RGB decimal format
  \definecolor{navyBlue}{rgb}{0.08,0.18,0.42}          % Darkened to a deeper navy blue
  \definecolor{petalColor}{rgb}{0.96,0.97,0.98}        % Clean, solid off-white
  \definecolor{softOrangeYellow}{rgb}{1.0,0.82,0.20}   % Soft yellow with a touch of orange
  \definecolor{brightYellow}{rgb}{1.0,0.92,0.0}        % Bright yellow for the border

  % 1. Darker Navy Blue Background Square
  \fill[navyBlue] (-2.2, -2.2) rectangle (2.2, 2.2);

  % 2. The 12-Petal Short and Fat Rose Curve
  \fill[
    petalColor,
    samples=400,
    domain=0:360,
    variable=\t
  ]
    plot (\t : {1.5 * pow(abs(cos(6 * \t)), 0.4)});

  % 3. Smaller Round Yellow Center Core with a Thinner Bright Yellow Outline
  % Line width reduced from 1.5pt to 0.8pt for a thinner border
  \filldraw[fill=softOrangeYellow, draw=brightYellow, line width=0.8pt] (0,0) circle (0.25);
\end{tikzpicture}
\xmalt{}
\end{center}


\end{document}