\documentclass{ximera}[font=12pt]

%Next 2 lines create spaces between items in itemize and enumerate
\setitemize{itemsep=1.5ex}
\setenumerate{itemsep=1.5ex}

%Creates space between paragraphs - Maybe???
\setlength{\parskip}{\medskipamount}


\title{Adding Integers: Different Signs}   % Use xmlatex all -s to compile when SERVE refuses to work.
\author{Kelly Stady}
\license{CC: 0}         % replace with an appropriate license, or set it in xmPreamble

\begin{document}

\begin{abstract}

\end{abstract}

\maketitle
\label{xim:AddIntDiffSActivity}


\section*{Adding Integers with Different Signs}

When adding integers with different signs, we can think of the process as a
\textbf{battle between positive and negative numbers}, where the winner is the sign of
the answer.


\textbf{Battle 1:}  \ If you have \$30 and spend \$20 to see a movie, your net balance is \$10. 

\[ 30 + (-20) = 10 \]

\textbf{The answer is positive, so the positive team win!}


\textbf{Battle 2:} \ If you have \$30 and a credit card debt of \$100, your net balance is a debt of \$70.


\[ 30 + (-100) = -70 \]

\textbf{The answer is negative, so the negative team win!}


Why did the positive team win the first battle and lose the second battle?

%The debt in the second battle was greater than the asset, so the positives lost that one.

In any number battle, the winner is always the team with the \textbf{greatest strength}, regardless of its sign.  Mathematically, we find the \textbf{absolute value} of a number to determine its \textbf{strength} or \textbf{size}.  Recall that the absolute value of a number is its distance from 0 on a number line.  Because it represents distance, it is \textbf{always positive}.

\textbf{Battle 1:} \ $30 + (-20) = 10$
\begin{itemize}
    \item The strength of the \$30 asset is $|30| = 30$.
    \item The strength of the \$20 debt is $|-20| = 20$.
\end{itemize}


Since $30$ is greater than $20,$ the \textbf{positive} team was stronger and \textbf{won} the battle.


\textbf{Battle 2:} \ $30 + (-100) = -70$
\begin{itemize}
    \item The strength of the \$30 asset is $|30| = 30$.
    \item The strength of the \$100 debt is $|-100| = 100$.
\end{itemize}

Since $30$ is smaller than $100,$ the \textbf{negative} team was stronger and \textbf{won} the battle.

We are now ready for a formal mathematical rule that can be applied to any problem, not just those involving money.  
As we noticed in our earlier examples, adding numbers with different signs is a bit of a paradox: to find the sum, 
we actually have to subtract.  From the number battles, we know that the sign of the answer depends on which number 
has the greatest strength. To turn these observations into a formal rule, we will use absolute value.


\begin{procedure}[\textbf{Steps for Adding Integers with Different Signs}]


\begin{enumerate}
    \item Find the \textbf{absolute value} or \textbf{strength} of each number.

     \item Subtract the smaller absolute value from the larger absolute value.

     \item Write the answer with the sign of the number with the \textbf{larger absolute value}.
\end{enumerate}
\end{procedure}

\begin{warning}
  % A small orange exclamation triangle
  \begin{tikzpicture}[baseline=(excl.base)]
    \node[draw=orange, thick, fill=orange!20, regular polygon, regular polygon sides=3, inner sep=1pt] (excl) {\textbf{!}};
  \end{tikzpicture}
  \quad % Adds a little horizontal space after the icon
  \textbf{Caution:} Always check if any orange-blue pairs can be removed before finalizing your sum.
\end{warning}



\begin{example} Let's use our three-step process to find the sum $-50 + 30.$

\begin{description}[style=multiline, leftmargin=5em, font=\bfseries]
    \item[Step 1] \textbf{Find the Absolute Values (Strengths):} \\
    First, we determine the size of each number.
    \[ |-50| = 50 \quad \text{and} \quad |30| = 30 \]

    \item[Step 2] \textbf{Subtract the Absolute Values (The Battle):} \\
    Since the signs are different, the numbers ``battle" by finding the difference between their strengths.
    \[ 50 - 30 = 20 \]

    \item[Step 3] \textbf{Declare the Winner (Determine the Sign):} \\
    Now we compare the absolute values from Step 1. Since $50$ is greater than $30,$ the negative team is stronger and wins the battle!
    \[ -50 + 30 = -20 \]
\end{description}

\end{example}


\bigskip

\begin{problem}
    Apply the steps to find the sum $15 + (-40).$

    ~ \\[3em]

    \textbf{Step 1:} Find the \textbf{absolute values} (strengths):
    \[ |15| = \answer{15} \quad \text{and} \quad |-40| = \answer{40} \]

    \begin{problem}
    \textbf{Step 2:} \textbf{Subtract} the smaller absolute value from the larger one (the battle):
    \[ 40 - 15 = \answer{25} \]


\begin{problem}
    \textbf{Step 3:} Declare the winner!
    Compare the absolute values: is $|15|$ or $|-40|$ larger?

    \begin{multipleChoice}
        \choice{The answer is positive, since $|15|$ is larger.}
        \choice[correct]{The answer is negative, since $|-40|$ is larger.}
    \end{multipleChoice}


\begin{problem}
    \textbf{Final Answer:} $15 + (-40) = \answer{-25}$

    \begin{feedback}[correct]
        Excellent! You found the strengths, fought the battle and declared the negative team the winner.
    \end{feedback}

\end{problem}

\end{problem}

\end{problem}
\end{problem}

\begin{problem}
Find the following sums.

\begin{enumerate}
\item
\end{enumerate}

\end{problem}

\end{document}