\documentclass{ximera}[font=12pt]

%Next 2 lines create spaces between items in itemize and enumerate
\setitemize{itemsep=1.5ex}
\setenumerate{itemsep=1.5ex}

%Creates space between paragraphs
\setlength{\parskip}{\medskipamount}


\title{Adding Integers}   % Use xmlatex all -s to compile when SERVE refuses to work.
\author{Kelly Stady}
\license{CC: 0}         % replace with an appropriate license, or set it in xmPreamble

\begin{document}   %%%%%%  ERROR when I try to include tikzlings!  ewFolder/ExponentsPowers10Activity.tex: LaTeX Error: File TikZlings.sty' not found  % https://ctan.org/pkg/tikzlings


\begin{abstract}

An introduction to adding integers.

\end{abstract}

\maketitle
\label{xim:AddIntroActivity}


\section*{The Set of Integers}

In this lesson, you will learn how to add numbers that belong to the set of integers.  The set of \textbf{integers} consists of the whole numbers along with the negative counting numbers.

\begin{description}[itemsep=1ex]  % an ex is the height of a lowercase x
    \item[\textbf{Counting Numbers}] $\{ 1, 2, 3, \ldots \}$
    \item[\textbf{Whole Numbers}] $\{0, 1, 2, 3, \ldots \}$
    \item[\textbf{Integers}] $\{ \ldots, -5, -4, -3, -2, -1, 0, 1, 2, 3, 4, 5, \ldots \}$
\end{description}

\Large{\textbf{Extra Insight}}

\normalsize
Although negative numbers were understood in China more than 2000 years ago and later in India, they were not immediately accepted in Europe.
For many centuries, European mathematicians were uncomfortable with the idea of numbers less than zero and sometimes referred to them  as ``false" or
``absurd" numbers.  However, during the 16th and 17th centuries, as algebra and calculus developed, the importance of negative numbers became undeniable.
Today, negative numbers are an essential part of mathematics and are commonly used to represent quantities such as losses, debts, temperatures below zero and
elevations below sea level.



\section*{Adding Integers}


To understand how to add integers, we will begin by using the Chinese Red-Black Rule and then a number line to visualize addition.  Then we'll use the concepts of assets and debts to understand addition in a real-world context.  These approaches will lead us to two rules: one for adding numbers with the same sign and another for adding numbers with different signs.


\subsection*{Adding Integers Using The Red-Black Rule}

Negative numbers were first used systematically in ancient China. In \emph{The Nine Chapters on the Mathematical Art} (written between about 200 BC and 100 AD), Chinese mathematicians used counting rods to represent numbers.  Red rods were used for positive numbers and black rods for negative numbers.  For example, the number 3 was
represented by three red rods and the number $-2$ was represented by two black rods.  These rods were used as a calculator of sorts for financial transactions involving profits and losses and to solve simple equations.  The Red-Black Rule is described below.

\begin{enumerate}
\item A red rod represents $1$ and a black rod $-1.$

\item To find the total of profits (red) and losses (black), place rods on a board to represent each number.

\item Since a profit and an equal loss result in 0, gather all pairs of red and black rods and remove them.

\item The total is what remains.
\end{enumerate}

\medskip

Since red and black can be difficult for some people to distinguish, we will use \textcolor{orange!85!black}{\textbf{orange rods}} to represent \textcolor{orange!85!black}{\textbf{positive numbers}} and \textcolor{blue!70!black}{\textbf{blue rods}} to represent \textcolor{blue!70!black}{\textbf{negative numbers}}.
We will now refer to this rule as the Orange-Blue Rule.

\bigskip

\textbf{Example} \ To find the sum $-5+3,$ begin with \textcolor{blue!70!black}{\textbf{five blue rods}} to represent the number $-5$ and \textcolor{orange!85!black}{\textbf{three orange rods}} to represent the number $3.$  Next remove all pairs of orange and blue rods.  Since two blue rods remain, the answer is $-2.$


\bigskip


\begin{tikzpicture}[scale=0.75, font=\sffamily]
% Define Accessible Colors
\colorlet{posColor}{orange!85!black}  % 85% orange, 15% black
\colorlet{negColor}{blue!70!black}

% --- SECTION 1: THE SETUP ---
    \begin{scope}[shift={(-5,0)}]
        \node[font=\bfseries\small, black!80, anchor=west] at (-0.3, 2.4) {Setup};

        % Value: -5
        \node[font=\bfseries, negColor, anchor=east] at (-0.2, 1.0) {$-5$};
        \foreach \x in {0, 0.4, 0.8, 1.2, 1.6} {
            \draw[line width=2.5pt, negColor, cap=round] (\x, 0.5) -- (\x, 1.5);
            %\node[negColor, font=\tiny\bfseries] at (\x, 1.7) {$-$};  % - Signs above Bars
        }

        % Value: +3
        \node[font=\bfseries, posColor, anchor=east] at (-0.2, -0.3) {$3$};
        \foreach \x in {0, 0.4, 0.8} {
            \draw[line width=2.5pt, posColor, cap=round] (\x, -0.8) -- (\x, 0.2);
            %\node[posColor, font=\tiny\bfseries] at (\x, -1.0) {$+$}; % + Signs above Bars
        }
    \end{scope}

    % --- SECTION 2: Remove Pairs ---
    \begin{scope}[shift={(0,0)}]
        \node[font=\bfseries\small] at (0.75, 2.4) {Remove Pairs};

        % The 3 Pairs that will cancel
        \foreach \x in {0, 0.4, 0.8} {
            \draw[line width=2.5pt, negColor, cap=round] (\x, 0.5) -- (\x, 1.5);
            \draw[line width=2.5pt, posColor, cap=round] (\x, -0.8) -- (\x, 0.2);
        }

        % CANCELLATION BOX
        \draw[dashed, black!40, line width=1pt, rounded corners=5pt] (-0.25, -1.1) rectangle (1.05, 1.9);

        % THE LEFTOVER BARS
        \foreach \x in {1.5, 1.9} {
            \draw[line width=2.5pt, negColor, cap=round] (\x, 0.5) -- (\x, 1.5);
            %\node[negColor, font=\tiny\bfseries] at (\x, 1.7) {$-$};
        }

        % Transition arrow
        \draw[->, ultra thick, black!30] (3.2, 0.3) -- (4.2, 0.3);
    \end{scope}

    % --- SECTION 3: THE RESULT ---
    \begin{scope}[shift={(5.5,0)}]
        \node[font=\bfseries\small] at (0.2, 2.4) {Result};

        % Only the 2 surviving Negative Rods
        \foreach \x in {0, 0.4} {
            \draw[line width=3pt, negColor, cap=round] (\x, 0.3) -- (\x, 1.3);
            %\node[negColor, font=\tiny\bfseries] at (\x, 1.5) {$-$};
        }

        % The Result Label
        \node[anchor=north, negColor, font=\bfseries] at (0.1, -0.2) {$-2$};
    \end{scope}

    % Bottom equation
    \node[anchor=north, black!80, font=\bfseries] at (-5, -2) {$-5 + 3 = -2$};

\end{tikzpicture}

\bigskip





\end{document}